Concurrent Programming with CXL

### 2020-06-04 book/paper.tex
Latest

The final na\"{\i}ve solution is based on a variable called \texttt{turn}
that alternates between 0 and 1.  When \texttt{turn = $i$}, process~$i$ can
enter the critical section, while process $1-i$ has to wait.  When done,
process~$i$ sets \texttt{turn} to $1-i$ to give the other process an
opportunity to enter.
An invariant of this solution is that while process~$i$ is in the critical
section, \texttt{turn} == $i$.
(If you do not believe that, replace the assertion at the critical
section with \texttt{assert turn == self} and run CXL.)
Since \texttt{turn} cannot be 0 and 1 at
the same time, mutual exclusion is satisfied.
The solution also has the nice property that
processes 0 and 1 alternate entering the critical section.
The problem with the solution is
that it also violates the \emph{progress} property:
if process~$i$ terminates instead of entering the critical section when it
is process~$i$'s turn, process $1-i$ ends up waiting indefinitely for its
turn.

@@

In 1981, Peterson came up with a beautiful solution to the mutual exclusion
problem, now known as ``Peterson's Algorithm''~\cite{Peterson81}.
The algorithm is an amalgam of the (incorrect) algorithms in
Figures~\ref{fig:upflags} and~\ref{fig:upturn}, and is presented
in Figure~\ref{fig:peterson}.
A process first indicates its interest in entering the critical section
by setting its flag.
It then politely gives way to the other process should it also want to
enter the critical section---if both do so at the same time one will
win because writes to memory in CXL are atomic.
The process continues to be polite, waiting in a \texttt{while} loop
until either the other process is no where near the critical section
or has given way.
Running the algorithm with CXL shows that it satisfies both mutual
exclusion and progress.

@@

A good way to prove mutual exclusion is through induction on
the number of steps.
In such a proof, we need a so-called
\emph{inductive invariant} $I$, which is some predicate that holds
over the state.  We need $I$ to satisfy the following:
\begin{itemize}
\item (ME1) $I$ should imply mutual exclusion.
\item (ME2) $I$ should hold in the initial state.
\item (ME3) If $I$ holds over some state $s$, and some process in the
state takes a step, then $I$ should also hold in the resulting state.
\end{itemize}

>>>>

The final na\"{\i}ve solution is based on a variable called \texttt{turn}
that alternates between 0 and 1.  When \texttt{turn = $i$}, process~$i$ can
enter the critical section, while process $1-i$ has to wait.  When done,
process~$i$ sets \texttt{turn} to $1-i$ to give the other process an
opportunity to enter.
An invariant of this solution is that while process~$i$ is in the critical
section, \texttt{turn} == $i$.
Since \texttt{turn} cannot be 0 and 1 at
the same time, mutual exclusion is satisfied.
The solution also has the nice property that
processes 0 and 1 alternate entering the critical section.
The problem with the solution is
that it also violates the \emph{progress} property:
if process~$i$ terminates instead of entering the critical section when it
is process~$i$'s turn, process $1-i$ ends up waiting indefinitely for its
turn.

@@

In 1981, Peterson came up with a beautiful solution to the mutual exclusion
problem, now known as ``Peterson's Algorithm''~\cite{Peterson81}.
The algorithm is an amalgam of the (incorrect) algorithms in
Figures~\ref{fig:upflags} and~\ref{fig:upturn}, and is presented
in Figure~\ref{fig:peterson}.
A process first indicates its interest in entering the critical section
by setting its flag.
It then politely gives way to the other process should it also want to
enter the critical section---if both do so at the same time one will
win because writes to memory in CXL are atomic.
The process continues to be polite, waiting in a \texttt{while} loop
until either the other process is nowhere near the critical section
or has given way.
Running the algorithm with CXL shows that it satisfies both mutual
exclusion and progress.

@@

A good way to prove mutual exclusion is through induction on
the number of steps.
In such a proof, we need a so-called
\emph{inductive invariant} $I$, which is some predicate that holds
over the state.  An inductive invariant $I$, by definition, satisfies the following:
\begin{itemize}
\item (ME1) $I$ implies mutual exclusion.
\item (ME2) $I$ hold in the initial state.
\item (ME3) If $I$ holds over some state $s$, and some process in the
state takes a step, then $I$ also holds in the resulting state.
\end{itemize}
